\documentclass[11pt]{article}

\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{enumitem}
\setlist{nosep, leftmargin=*, topsep=2pt, itemsep=2pt}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue!50!black, urlcolor=blue!50!black,
            pdftitle={Outline -- Rewiring the market time clock in Spanish electricity markets},
            pdfauthor={Pablo Paramio Perez}}
\usepackage{microtype}
\linespread{1.10}

\title{Outline\\
\large Rewiring the market time clock in Spanish electricity markets}
\author{Pablo Paramio P\'erez}
\date{\today}

\begin{document}
\maketitle

\section*{RAP}

\noindent\textbf{Positioning.} How shorter trading intervals reshape
wholesale electricity bidding and prices has been studied for Germany
(M\"arkle-Hu\ss, Feuerriegel and Neumann, 2017, on the 2014 EPEX
15-minute reform) and for Australia (Csereklyei and Khezr, 2024, on
the 2021 five-minute settlement reform), but the recent Spanish move
to 15-minute trading has not yet been evaluated. The Spanish case is
distinctive: the reform was rolled out in two cutovers (intraday in
March 2025, day-ahead in October 2025) and lands in the middle of a
nationwide blackout (28 April 2025) and the post-blackout reinforced
operation regime that has been expanding regulated redispatch since.

\noindent\textbf{Research Question.} How did Spain's switch from
hourly to 15-minute wholesale trading change the way generators bid,
how the auctions clear, and how the system as a whole settles? In
particular: who actually uses the new within-hour freedom, and where
does it bite? Does the reform amplify market power, or does it
compress it? Does it harm consumers? And how can the day-ahead
reform's effect be separated from the concurrent post-blackout
regulatory regime?

\noindent\textbf{Answer.} The 15-minute trading reform acts not as a
uniform technology shock but as a \emph{strategic-latitude reveal}:
its uptake maps to who has the economic latitude to use it
strategically. Dispatchable firms (combined-cycle gas, hydro,
pumped-hydro) widen their within-quarter bid ladders in
high-residual-demand hours (the morning ramp and evening peak), where
within-hour residual demand actually varies and rewards finer pricing;
flat hours barely move. Wind and solar aggregators also use the new
grid, but for a different reason --- imbalance-penalty avoidance under
the December 2024 settlement reform --- which applies equally in every
hour, and so leaves no critical-flat differential. On the auction
side, intraday clearing prices drop sharply under the March 2025
reform with day-ahead co-moving down via cross-market anticipation;
the day-ahead combined-cycle gas auction-cleared volume scales up by
about $59$~GWh/day under the October 2025 reform in critical hours;
the day-ahead/intraday wedge keeps its mean at zero but its within-day
standard deviation jumps in critical hours under the intraday reform.
Granularity does not amplify market power (prices fall, the wedge
level is unchanged) and does not harm consumers (the granularity
reforms compress the real-time redispatch and reserve activations REE
has to call after the auction; the post-blackout reforzada cost rise
is a separate mechanism that happens to share calendar time). The
interpretive frame is that the reform restructures the whole
sequential-market architecture even where it does not touch the
auction directly: under the intraday cutover, combined-cycle gas
widens its day-ahead bid-shape \emph{in anticipation} of finer
intraday re-trading.

\section*{Questions Provoked by the RAP}

\begin{enumerate}
  \item What does the move from hourly to 15-minute trading actually
        change in the wholesale market, and how is the answer
        different for dispatchable units vs.\ renewables?
  \item How can the day-ahead reform's effect be separated from the
        concurrent post-blackout reinforced-operation regime?
  \item How does bid widening translate into the day-ahead
        combined-cycle gas cleared-volume scale-up in critical hours?
  \item Does the reform amplify market power, and does it harm
        consumers, once both the wholesale auction and the system-cost
        cascade are accounted for?
  \item Could the March 2025 intraday reform have contributed to the
        macro-conditions that made the April 2025 blackout possible?
\end{enumerate}

\section{Introduction}

The two-step Spanish 15-minute trading reform --- intraday auctions in
March 2025, day-ahead market in October 2025 --- arrives in an
institutional setting that has been reshaped by a nationwide blackout
(28 April 2025) and a continuous reinforced-operation
(\emph{reforzada}) regime that has expanded regulated redispatch ever
since. We ask whether the reform changed wholesale bidding, prices and
cleared quantities; whether it amplifies market power or compresses
it; whether it harms consumers; and how the day-ahead cutover can be
identified given the concurrent regulatory regime. The answer is that
the reform is a strategic-latitude reveal: dispatchable firms widen
their within-quarter bid spreads where within-hour residual demand
actually varies, renewables use the finer grid uniformly to manage
imbalance, the intraday auction price drops by $\sim 80$~EUR/MWh
placebo-net under the March 2025 reform, the day-ahead combined-cycle
gas cleared volume scales up by $\sim 59$~GWh/day under the October
2025 reform in critical hours, and a wedge-SD jump shows that
within-hour bid behavior also restructures the day-ahead/intraday
relationship even though the wedge's level stays at zero. The reform
does not amplify market power or harm consumers: wholesale prices
fall, the wedge level stays at zero, the up-direction adjustment-cost
channels REE has to call shrink; the post-blackout cost rise is a
separate, reforzada-driven effect that overlaps in calendar time but
operates through a different mechanism (Fase~I redispatch).

\textit{Questions answered by the introduction.}
\begin{enumerate}
  \item What is the puzzle, given the German and Australian
        precedents and the recent Spanish reform?
  \item What is the paper's precise answer, and what unifying
        mechanism does it propose?
  \item How does the rest of the paper establish the result, and
        what does the institutional setting force us to control for?
\end{enumerate}

\section{Institutional Setting: Sequential Markets, Three Reforms, and a Blackout}

Spanish wholesale clearing is sequential: a day-ahead auction first
(OMIE), then three intraday auctions, then continuous trading
(XBID/SIDC) and balancing. The 15-minute trading reform was rolled out
in three steps that interact in sequence: a 15-minute settlement
reform in December 2024 (ISP15), the intraday cutover in March 2025
(ID15), and the day-ahead cutover in October 2025 (DA15). The blackout
of 28 April 2025 switches the system into a continuous
reinforced-operation regime that overlaps only with the day-ahead
cutover, not with the intraday one. Granular pricing only has
economic content where within-hour residual demand actually varies, so
we partition clock-hours into ``critical'' (the morning ramp and the
evening peak, where the standard deviation of 15-minute residual
demand is high) and ``flat'' (overnight, the within-day control).

\textit{Questions answered by this section.}
\begin{enumerate}
  \item How do day-ahead and intraday auctions relate, and where does
        regulated redispatch sit on top of them?
  \item What changed mechanically on 11 December 2024, 19 March 2025
        and 1 October 2025, and what did not?
  \item Why is the critical/flat partition the right place to look
        for a granularity effect, and why is the partition itself
        identification-relevant?
\end{enumerate}

\subsection{Three reforms and a blackout (Dec 2024, Mar 2025, Apr 2025, Oct 2025)}
The December 2024 settlement reform (ISP15) moves the imbalance
settlement period from 60 to 15 minutes but leaves the wholesale
contract grid unchanged. The intraday cutover (19 March 2025) moves
all three intraday auctions and the continuous intraday market to
15-minute trading; the day-ahead cutover (1 October 2025) does the
same for the day-ahead market. The 28 April 2025 blackout triggers a
continuous reinforced-operation regime that overlaps only with the
day-ahead cutover.

\subsection{Where granular pricing has economic content: the critical/flat partition}
A clock-hour is ``critical'' if the standard deviation of 15-minute
residual demand exceeds the cross-hour median, and ``flat'' otherwise;
the flat hours serve as the within-day control group for the bid-shape
difference-in-differences. Critical here is the morning-ramp $\cup$
evening-ramp union; midday (solar-saturated) is excluded from the
main specifications and reserved as a falsification.

\section{Data: OMIE Bid-Level Offers, ENTSO-E Weather and Load, ESIOS Redispatch}

We use OMIE per-curve sell offers for every unit, date and period in
the day-ahead and intraday-auction markets from 2022 to 2026, aligned
to the corresponding OMIE clearing prices and complemented by 15-minute
ENTSO-E wind, solar and load and by ESIOS Fase~I/II and aFRR
regulated-redispatch volumes and prices. Bid tranches are restricted
to a window-and-market-specific band around each clearing price (set
to the $p_{90}-p_{50}$ spread of the MCP distribution in the analysis
window) to isolate the bids that plausibly compete with the marginal
price; the critical/flat partition is calibrated on
post-15-minute-trading data.

\textit{Questions answered by this section.}
\begin{enumerate}
  \item What is a ``curve'' here, and what bid-shape measures are
        defined on it?
  \item Why a window-specific bandwidth, and how sensitive are the
        conclusions to this choice?
  \item How are OMIE's three offer types (simple, MIC, block) used in
        the analysis, and why does the offer-type composition matter
        for the bid-shape DiD?
\end{enumerate}

\subsection{Per-curve bid-shape outcomes and the in-band offered energy}
Each (unit, date, period) bid ladder yields two per-curve bid-shape
measures: the MW-weighted in-band standard deviation of bid prices
$\sigma_p$ and the effective tranche count $N_{\text{eff}}$. A
complementary aggregate-volume primitive is the in-band offered energy
$E^{\text{band}}$ (in MWh per class-hour, normalised so that the
quarter-count expansion under 15-minute trading does not mechanically
inflate the level).

\subsection{Calibrating the critical/flat partition (post-only)}
Date-fixed-effect regressions of within-hour dispersion on critical
and midday dummies, fit on post-15-minute-trading data, confirm that
critical hours are where within-hour bid-shape dispersion actually
lives across all three dispatchable techs.

\section{Empirical Strategy: Three Specifications for Three Economic Objects}

The reform's effect lives in three economic objects --- bid shape,
in-band offered energy, and auction-clearing outcomes --- and each
calls for a different specification. The bid-shape DiD identifies the
critical-minus-flat shift across each cutover under within-day
parallel trends, with unit fixed effects and date-clustered standard
errors. The auction-clearing and aggregate-volume outcomes use a
Bayesian counterfactual model (BSTS / CausalImpact) with wind, solar
and gas covariates plus a 2024 same-calendar placebo, mirroring
M\"arkle-Hu\ss\ et al.\ (2017). Identification of the day-ahead
cutover requires holding the post-blackout reinforced-operation
regime constant: the day-ahead pre-window begins on 28 April 2025
(the day of the blackout) and ends on 30 September 2025, so both pre
and post sit inside the \emph{reforzada} period.

\textit{Questions answered by this section.}
\begin{enumerate}
  \item Why difference-in-differences for bid shape and a Bayesian
        counterfactual for prices and quantities?
  \item What identifying assumption does the day-ahead pre-window
        choice rely on, and what would break with a naive long pre?
  \item How does the 2024 same-calendar placebo correct for residual
        seasonal trend?
\end{enumerate}

\subsection{Spec A: Bayesian counterfactual for prices, cleared MW, and the day-ahead/intraday wedge SD}
Daily outcomes (clearing prices, per-tech cleared GWh, within-day
wedge SD) are modelled with a state-space local-linear trend, weekly
seasonality and wind/solar/gas regressors trained pre-cutover and
extrapolated forward; the post-cutover gap from the counterfactual is
the treatment effect, with a 2024 same-calendar placebo subtracted
for the joint $95\%$ test.

\subsection{Spec B: per-hour-class BSTS on in-band offered energy}
The same Spec~A architecture is run independently per (tech, market,
hour-class) cell on $E^{\text{band}}$, the in-band offered energy
normalised to MWh per class-hour. The readable summary is the real
posterior and the 2024 placebo posterior reported side by side per
hour-class; the reform-attributable shift is the gap.

\subsection{Spec C: per-curve DiD for bid shape}
Per-curve DiD on $\sigma_p$ and $N_{\text{eff}}$ with unit fixed
effects and the critical/flat contrast. The DiD is the right tool
here precisely because the granularity reform's effect on bid shape
is structurally a critical-hours effect: differentiated bidding only
pays off where within-hour residual demand actually varies. The
critical $-$ flat differential $\theta$ \emph{is} the
granularity-enabled discrimination, not a residual being netted out.

\section{Results}

The empirical findings sort into five clusters, layered from bidding
behavior up to auction-clearing outcomes and cross-market spillovers.
\emph{Bid-shape widening} shows up under both cutovers and
concentrates in critical hours where the steep residual-demand curve
rewards within-quarter discrimination. \emph{In-band offered energy}
migrates toward the newly-granular market under both reforms, with
the morning ramp leading under the day-ahead cutover. On the
auction-clearing side, the intraday reform drops both intraday and
day-ahead clearing prices through a cross-market anticipation
channel; the day-ahead reform scales up combined-cycle gas
auction-cleared volume in critical hours by $\sim 59$~GWh/day; the
day-ahead/intraday wedge keeps its mean at zero but its within-day
standard deviation jumps under the intraday cutover. Finally, a
\emph{cross-market substitution} pattern --- combined-cycle gas
widens its day-ahead bid-shape \emph{in anticipation} of finer
intraday re-trading under the intraday cutover, and tightens its
intraday bid-shape under the day-ahead cutover --- shows that the
reform restructures the whole sequential-market architecture, not
just the auction it directly touches. The \emph{operational vs.\
strategic} distinction explains why wind shows null DiD: wind
aggregators face operational (imbalance-driven) incentives that apply
equally in every hour and so leave no critical-flat differential,
while dispatchable firms face strategic (rent-maximising) incentives
that concentrate in critical hours.

\textit{Questions answered by this section.}
\begin{enumerate}
  \item Which findings survive both a within-pre placebo and a 2024
        same-calendar placebo, and which need the joint $95\%$
        bracket?
  \item Why does the bid widening translate into more day-ahead
        cleared volume but not into a day-ahead wholesale price
        change?
  \item How does the wedge-SD jump reconcile with the wedge-level
        null --- and what does that say about the within-hour
        restructuring?
\end{enumerate}

\subsection{Bid-shape widens in critical hours --- both reforms (Spec C, Finding 1)}
Combined-cycle gas, hydro and pumped-hydro significantly widen their
within-quarter bid-price spread $\sigma_p$ and effective tranche
count $N_{\text{eff}}$ in critical hours after each reform; flat
hours barely move. Wind widens with magnitudes an order of magnitude
smaller --- the operational vs.\ strategic distinction is what
collapses its critical-flat differential, not pure price-taking.

\subsection{In-band offered energy migrates into the granular market --- both reforms (Spec B, Finding 2)}
The dispatchable bid-side migration shows up directly as MWh shifting
in-band toward the newly-granular market: under the day-ahead reform
the MWh move into the day-ahead auctions (morning ramp dominates),
under the intraday reform into the intraday auctions.

\subsection{Auction-clearing outcomes (Spec A, Finding 3)}
The intraday reform drops both intraday and day-ahead clearing prices
by $\sim 80$~EUR/MWh placebo-net (the day-ahead drop is the
rational-anticipation channel into day-ahead); the day-ahead reform
scales up combined-cycle gas day-ahead cleared volume in critical
hours by $\sim 59$~GWh/day; renewables reallocate cleared MW toward
the granular market under both reforms (ISP15-driven imbalance
avoidance); the intraday reform leaves the per-clock-hour day-ahead
$-$ intraday wedge level at zero but jumps its within-day SD in
critical hours.

\subsection{Cross-market substitution: the reform restructures the whole sequential-market architecture (Spec C, Finding 4)}
Under the intraday reform, combined-cycle gas widens its day-ahead
$\sigma_p$ \emph{in anticipation} of finer intraday re-trading;
under the day-ahead reform, combined-cycle gas tightens its intraday
$\sigma_p$, pulling structured bidding into the now-granular
day-ahead. The granularity instrument rearranges the whole sequential
market, not just the auction it directly touches.

\subsection{Operational vs.\ strategic incentives explain the wind/solar null DiD (Finding 5)}
Wind and solar aggregators bid at quarter granularity in every hour
to manage imbalance exposure under the December 2024 settlement
reform; that engagement is real but uniform across hour-classes, so
the critical-flat DiD collapses for the renewables even though they
are not pure price-takers. Dispatchable firms have no comparable
operational pressure and engage strategically only where
within-hour residual demand actually varies.

\section{Robustness}

The bid-shape and auction-clearing findings survive the standard
checks: a within-pre midpoint placebo (the intraday-reform bid-shape
findings are clean; the day-ahead-reform combined-cycle gas
$\sigma_p$ widening cleans up after de-trending), a midday-vs-flat
hour-class falsification (sign-stable, smaller magnitudes), a
cross-border-flow control on the intraday $\sigma_p$ (essentially
unchanged), a per-CCGT-firm decomposition (the widening is driven by
two of the four firms, the two most active in the day-ahead
competing tier), a bandwidth sensitivity check (the $p_{90}$ choice
isolates the competing tier from the parked scarcity tier; wider
bandwidths contaminate the measure), a per-IDA-session BSTS for the
intraday price drop (session-symmetric), and an \emph{offer-type
composition check}: the combined-cycle gas in-band offer mix does
shift across the day-ahead reform (simple-offer share drops $29\%
\to 22\%$, block share grows $30\% \to 35\%$), but in the direction
that should dampen the pooled $\sigma_p$ widening; restricting to
simple offers only strengthens the effect from $1.25^{**}$ to
$3.53^{***}$, so the pooled headline is a conservative read of the
true within-tranche strategic engagement.

\textit{Questions answered by this section.}
\begin{enumerate}
  \item Which findings rely on a single robustness leg, and which on
        multiple?
  \item Why is the offer-type composition shift not a confounder for
        the bid-shape DiD?
  \item Why is \emph{reforzada} a level confound rather than a
        within-hour confound?
\end{enumerate}

\section{Mechanism: Where the Strategic Margin Lives and Where Granularity Efficiency Shows Up}

Two market-structure facts frame the bidding evidence. First, each
Big-4 combined-cycle fleet bids a two-tier day-ahead curve --- a
``competing'' tier near marginal cost and a ``parked'' tier above
the clearing-price ceiling --- and a single competitive tier at
marginal cost in the intraday auction, indicating that the strategic
margin is collected through regulated-redispatch recall, not through
wholesale price-setting. Second, wholesale concentration (HHI $\sim$
1,100) is moderate while ancillary-market concentration (HHI $>$
4,000) is high, so market power migrated from wholesale to ancillary
services as renewables eroded the wholesale margin. On the cost
side, the granularity reforms compress the up-direction
adjustment-cost channels REE has to call after the auction
(real-time redispatch and aFRR reserve activations both shrink
across the reform sequence) --- the efficiency channel of 15-minute
trading. The post-blackout reforzada cost rise concentrates on
combined-cycle gas (Fase~I up balloons; other technologies are flat)
and is a separate policy effect that happens to overlap calendar-wise
with the day-ahead cutover but operates on a different mechanism.

\textit{Questions answered by this section.}
\begin{enumerate}
  \item What is the geometry of the two-tier combined-cycle gas
        day-ahead curve, and why is the parked tier never recalled
        by the wholesale auction?
  \item Why is the absence of wholesale price effects under the
        day-ahead reform consistent with the cleared-volume scale-up,
        the cross-market substitution, and the migration of strategic
        margin to ancillary services?
  \item How do the granularity efficiency gains and the
        reforzada-driven cost rise interact in calendar time, and
        how do we keep them separate in the analysis?
\end{enumerate}

\subsection{Two-tier combined-cycle gas day-ahead curve and the day-ahead/intraday contrast}
The same fleets bid a single competitive tier with negligible
withholding in the intraday auction, after Fase~I has cleared ---
indicating that the parked tier is a non-strategic placeholder
rather than a Fase~I bid, and the strategic margin is collected
through a separate restriction offer submitted to REE.

\subsection{Concentration migrated from wholesale to ancillary services}
Wholesale-market HHI hovers around 1,100 over 2016--2024 and falls
mid-day when solar peaks; ancillary-market HHI sits above 4,000
across the same window. The cleared-volume scale-up under the
day-ahead reform is the combined-cycle gas competing tier expanding
into the four quarters of each peak hour, scaling up the volume base
that feeds the redispatch recall.

\subsection{Granularity compresses the adjustment-cost channels --- the efficiency channel}
Finer trading-period resolution lets the auction itself absorb
within-hour variation that previously had to be cleaned up by
ex-post REE intervention. Real-time redispatch falls roughly
threefold across the reform sequence; aFRR follows. Solar
aggregators exploit granularity in midday hours by committing to
four quarter-level schedules that track cloud-cover variation,
instead of being locked into a single hourly commitment.

\subsection{Reforzada cost increase concentrates on combined-cycle gas --- a separate mechanism}
Decomposing weekly dispatch into its components, the combined-cycle
gas day-ahead-cleared layer collapses post-blackout while the
intraday $+$ REE real-time top-up layer balloons; other technologies
are flat. A granularity reform that touched all bid-side firms
equally would not produce a combined-cycle-gas-only intervention
pattern --- the reforzada cost rise is a separate policy effect from
the granularity reforms.

\section{Conclusion: Granularity Is a Strategic-Latitude Reveal}

A market-microstructure reform that simply increases trading-period
resolution acts not as a uniform technology shock but as a
\emph{strategic-latitude reveal}: it does not change marginal costs,
capacity, or who is in the market; it changes who has room to
discriminate within the hour and where, and the equilibrium response
traces exactly that --- strategic firms moving in strategic hours,
operational firms moving uniformly, both rippling across the
sequential markets through cross-market anticipation, and all of it
within (not against) the competitive clearing. The reform does not
amplify market power (wholesale prices fall, the wedge level is
unchanged) and does not harm consumers (the granularity reforms
compress the real-time adjustment-cost channels). The Fase~I cost
rise on the same calendar window is the post-blackout
reinforced-operation response, not the granularity reforms: the two
policies overlap in time but operate on different mechanisms.
Whether the March 2025 intraday reform was in some way responsible
for the April 2025 blackout is out of scope here --- the ENTSO-E
Expert Panel's final report does not list the reform as a
contributing factor; the CNMC report does flag the move to
quarter-hourly trading as one contributor to the abrupt voltage
variations observed in the months around the incident. Answering it
is beyond the scope of this thesis (it requires
electrical-engineering knowledge we do not have) but it points to a
strong research gap: the ENTSO-E report identifies the proximate
technical triggers but does not engage with whether the
market-microstructure reforms contributed to the macro-conditions
that made the system fragile enough for those triggers to cascade.

\textit{Questions answered by the conclusion.}
\begin{enumerate}
  \item What is the main empirical lesson, and what is the surviving
        level-shift finding?
  \item What is the interpretive frame, and what would tighten it
        from descriptive to identified?
  \item What extensions follow naturally (the ongoing
        functional-factor regression for the full curve, a
        firm-level decomposition, identifying the redispatch
        channel)?
\end{enumerate}

\end{document}
